---
title: "BenchHub Case Study: Spatial domain detection"
date: "`r BiocStyle::doc_date()`"
output: 
  BiocStyle::html_document:
    toc_float: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{BenchHub Case Study: Spatial domain detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning=FALSE
)
```

```{r, message=FALSE, warning=FALSE}
devtools::load_all()
library(BenchHub)
library(FNN)
library(igraph)
library(Matrix)
library(SingleCellExperiment)
library(readr)
library(dplyr)
library(tidyr)
```

```{r}
Sys.setenv(GITHUB_PAT = "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")
```


```{r}
tempCache <- tempdir()
ARI <- function(auxData, predicted) {
  result <- mclust::adjustedRandIndex(auxData, predicted)
  return(result)
}

NMI <- function(auxData, predicted) {
  result <- aricode::NMI(auxData, predicted)
  return(result)
}

PAS <- function(auxData, predicted) {
  # browser() 
  k <- 10        
  threshold <- 6 
  nn_result <- get.knn(predicted, k = k)
  abnormal_count <- sum(sapply(1:nrow(predicted), function(i) {
    neighbor_labels <- auxData[nn_result$nn.index[i, ]]
    sum(neighbor_labels != auxData[i]) >= threshold
  }))
  return(abnormal_count / nrow(predicted))
}

CHAOS <- function(auxData, predicted) {
  cluster_ids <- unique(auxData)
  cluster_ids <- cluster_ids[!is.na(cluster_ids)] 
  
  chaos_scores <- sapply(cluster_ids, function(cluster) {
    indices <- which(auxData == cluster)
    coords <- predicted[indices, , drop = FALSE]
    
    if (nrow(coords) <= 1) return(0) 
    
    dist_mat <- as.matrix(dist(coords))
    
    g <- graph_from_adjacency_matrix(dist_mat, mode = "undirected", weighted = TRUE)
    mst <- mst(g)
    mst_sum <- sum(E(mst)$weight)
    normalized_mst <- mst_sum / nrow(coords)
    return(normalized_mst)
  })
  
  return(mean(chaos_scores))
}

data <- readRDS("~/Downloads/STARmap.rds")

# Mouse Primary cortex under BaristaSeq Protocol
# Human Postmortem DLPFC under 10x Visium Protocol

# Mouse Preoptic hypothalamus under MERFISH Protocol
# Human Pancreatic ductal adenocarcinomas under ST Protocol
# Mouse Medial prefrontal cortex under STARmap Protocol


# Mouse embryo under seqFISH Protocol




trio <- BenchHub::Trio$new(data = data, cachePath = tempCache)

trio$addMetric("ARI", ARI)
trio$addMetric("NMI", NMI)
trio$addMetric("PAS", PAS)
trio$addMetric("CHAOS", CHAOS)

trio$addEvidence(
   name = "Accuracy", 
   evidence = setNames(colData(trio$data)$manual_anno,
                       colnames(trio$data)), 
   metrics = c("ARI", "NMI")
)

trio$addEvidence(
  name = "Continuity", evidence = as.matrix(colData(trio$data)[, c("row", "col")]), metrics = c("PAS", "CHAOS")
)

# trio$addEvidence(
#   name = "scalability", evidence = NA, metrics = c("time",  # "memory")
# )

BenchHub:::createGist("\\(x) NULL", 'test.R', 'test')

trio$writeCTD("MMPstarmap")



```



# Prepare metric

```{r, eval=FALSE}
tempCache <- tempdir()
ARI <- function(auxData, predicted) {
  result <- mclust::adjustedRandIndex(auxData, predicted)
  return(result)
}

NMI <- function(auxData, predicted) {
  result <- aricode::NMI(auxData, predicted)
  return(result)
}

PAS <- function(auxData, predicted) {
  # browser() 
  k <- 10        
  threshold <- 6 
  nn_result <- get.knn(predicted, k = k)
  abnormal_count <- sum(sapply(1:nrow(predicted), function(i) {
    neighbor_labels <- auxData[nn_result$nn.index[i, ]]
    sum(neighbor_labels != auxData[i]) >= threshold
  }))
  return(abnormal_count / nrow(predicted))
}

CHAOS <- function(auxData, predicted) {
  cluster_ids <- unique(auxData)
  cluster_ids <- cluster_ids[!is.na(cluster_ids)] 
  
  chaos_scores <- sapply(cluster_ids, function(cluster) {
    indices <- which(auxData == cluster)
    coords <- predicted[indices, , drop = FALSE]
    
    if (nrow(coords) <= 1) return(0) 
    
    dist_mat <- as.matrix(dist(coords))
    
    g <- graph_from_adjacency_matrix(dist_mat, mode = "undirected", weighted = TRUE)
    mst <- mst(g)
    mst_sum <- sum(E(mst)$weight)
    normalized_mst <- mst_sum / nrow(coords)
    return(normalized_mst)
  })
  
  return(mean(chaos_scores))
}



```

# Set up and save trio

```{r, eval=FALSE}
output_dir <- "/Users/cabiria/Desktop/manuscript/TrioR/vignettes/trio_objects"
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

datasets <- list(
  MOB = list(file = "~/Downloads/result/MOB.csv", id = "figshare:29142950/54794756"),
  PDAC = list(file = "~/Downloads/result/PDAC.csv", id = "figshare:29142950/54794762"),
  MGASTRULA = list(file = "~/Downloads/result/MGASTRULA.csv", id = "figshare:29142950/54794759"),
  STARmap = list(file = "~/Downloads/result/STARmap.csv", id = "figshare:29142950/54794771"),
  SMI = list(file = "~/Downloads/result/SMI.csv", id = "figshare:29142950/54794768"),
  seqfish = list(file = "~/Downloads/result/seqfish.csv", id = "figshare:29142950/54794765"),
  baristaSeq = list(file = "~/Downloads/result/baristaSeq.csv", id = "figshare:29142950/54794750") ,
  LIBD = list(file = "~/Downloads/result/LIBD.csv", id = "figshare:29142950/54794753"),
  MERFISH = list(file = "~/Downloads/result/MOB.csv", id = "figshare:29142950/54794747")
)

# Loop over datasets
for (name in names(datasets)) {
  # Load CSV and ID
  
  name <- "MOB"
  
  df <- read_csv(datasets[[name]]$file)
  df <- df[complete.cases(df), ]
  figshare_id <- datasets[[name]]$id
  
  # Create new Trio object
  trio <- Trio$new(figshare_id, cachePath = tempCache)
  
  # Add metrics
  trio$addMetric("ARI", ARI)
  trio$addMetric("NMI", NMI)
  trio$addMetric("PAS", PAS)
  trio$addMetric("CHAOS", CHAOS)
  
  # Add auxiliary data
  trio$addEvidence(
    name = "Accuracy", evidence = df$ground_truth, metrics = c("ARI", "NMI")
  )
  trio$addEvidence(
    name = "Continuity", evidence = as.matrix(df[, c("row", "col")]), metrics = c("PAS", "CHAOS")
  )
  
  saveRDS(trio, file = file.path(output_dir, paste0(name, "_trio_test.rds")))
  
}
```

```{r}
googlesheets4::gs4_auth()
trio$writeCTD(name = "MOBTEST")
```




# Load the trio and get evaluation result

```{r}
datasets <- list(
  MOB = list(file = "~/Downloads/result/MOB.csv", trio = "trio_objects/MOB_trio.rds"),
  PDAC = list(file = "~/Downloads/result/PDAC.csv", trio = "trio_objects/PDAC_trio.rds"),
  MGASTRULA = list(file = "~/Downloads/result/MGASTRULA.csv", trio = "trio_objects/MGASTRULA_trio.rds"),
  STARmap = list(file = "~/Downloads/result/STARmap.csv", trio = "trio_objects/STARmap_trio.rds"),
  # SMI = list(file = "~/Downloads/result/SMI.csv", trio = "~/trio_objects/SMI_trio.rds"),
  # seqfish = list(file = "~/Downloads/result/seqfish.csv", trio = "trio_objects/seqfish_trio.rds"),
  baristaSeq = list(file = "~/Downloads/result/baristaSeq.csv", trio = "trio_objects/baristaSeq_trio.rds"),
  LIBD = list(file = "~/Downloads/result/LIBD.csv", trio = "trio_objects/LIBD_trio.rds"),
  MERFISH = list(file = "~/Downloads/result/MOB.csv", trio = "trio_objects/MERFISH_trio.rds")
)

# Initialize final result
final_result <- NULL

# Methods to evaluate
methods <- c("BASS", "SpatialPCA", "leiden", "BayesSpace", "PRECAST", "DESC")


# Loop through datasets
for (name in names(datasets)) {
  cat("Evaluating:", name, "\n")

  # Load prediction data
  df <- read_csv(datasets[[name]]$file)
  df <- df[complete.cases(df), ]

  # Load the pre-saved Trio object
  trio <- readRDS(datasets[[name]]$trio)

  # Create evaluation input
   result <- trio$evaluate(list(
    BASS = list("Accuracy" = df$spatial_cluster_BASS) ,
    SpatialPCA = list("Accuracy" = df$spatial_cluster_SpatialPCA),
    leiden = list("Accuracy" = df$spatial_cluster_leiden),
    BayesSpace = list("Accuracy" = df$spatial_cluster_BayesSpace),
    PRECAST = list("Accuracy" = df$spatial_cluster_PRECAST),
    DESC = list("Accuracy" = df$spatial_cluster_DESC),
  
    BASS = list("Continuity" = df$spatial_cluster_BASS),
    SpatialPCA = list("Continuity" = df$spatial_cluster_SpatialPCA),
    leiden = list("Continuity" = df$spatial_cluster_leiden),
    BayesSpace = list("Continuity" = df$spatial_cluster_BayesSpace),
    PRECAST = list("Continuity" = df$spatial_cluster_PRECAST),
    DESC = list("Continuity" = df$spatial_cluster_DESC)
  ))

  # Append to final result
  final_result <- bind_rows(final_result, result)
}


# insert scability
scability <- read_csv("~/Downloads/benchmark_results.csv")

scability_long <- scability %>%
  pivot_longer(cols = c(time, memory), names_to = "metric", values_to = "result") %>%
  mutate(
    datasetID = "29142950/54794753",
    auxData = case_when(
      dataset == "LIBD_fixSpot_1k_gene1k" ~ 10,
      dataset == "LIBD_fixSpot_1k_gene5k" ~ 50,
      dataset == "LIBD_fixGene_2k_spot1k" ~ 20,
      dataset == "LIBD_fixGene_2k_spot3k" ~ 60,
      TRUE ~ NA_real_
    )
  ) %>%
  dplyr::select(datasetID, method, auxData, metric, result)

final_result <- rbind(final_result, scability_long)

knitr::kable(head(final_result, 10))
```

# Set up benchamrkInsight

```{r}
bmi <- benchmarkInsights$new(final_result)
bmi
```

# Get insight from case study

**Case Study: What is the overview of summary?**

```{r}
bmi$getHeatmap(bmi$evalSummary)
```

**Case Study: What is the correlation between auxData/metric/method?**

```{r}
p1 <- bmi$getCorplot(bmi$evalSummary, "method") + scale_fill_gradient2(
    low = "blue",       # for negative correlation
    mid = "white",      # for zero correlation
    high = "#B1886A",   # for positive correlation
    midpoint = 0,
    limits = c(0, 1),
    name = "Correlation"
  )+
  theme(legend.position = "bottom")
```

```{r}
p2 <- bmi$getScatterplot(bmi$evalSummary, c("ARI", "NMI")) +
  scale_color_manual(values = my_colors)
```

**Case Study: What is the time and memory trend?**

```{r}
my_colors <- c("#B9DAD3", "#CBD38A", "#DF7488", "#F9C49C", "#C99158", "#AD848C")

p3<- bmi$getLineplot(bmi$evalSummary, metricVariable = "memory") +
  scale_color_manual(values = my_colors)+
  theme(legend.position = "bottom")
```

```{r}
bmi$getLineplot(bmi$evalSummary, metricVariable = "time")
```


**Case Study: Which metric is most effective on the method?**

```{r}
bmi$getForestplot(bmi$evalSummary, "metric", "method")
```

**Case Study: How does method variability differ across datasets for a specific metric?**

```{r}
my_colors <- c("#B9DAD3", "#CBD38A", "#DF7488", "#F9C49C", "#C99158", "#AD848C", "#8C8C8C")

p4 <- bmi$getBoxplot(bmi$evalSummary, metricVariable = "ARI", auxDataVariable = "Accuracy") +
  scale_fill_manual(values = my_colors)+
  theme(legend.position = "bottom")

bmi$getBoxplot(bmi$evalSummary, metricVariable = "PAS", auxDataVariable = "Continuity")
```

```{r}
combined <- ggarrange(p1, p2, p3, p4, nrow = 1, ncol = 4)

ggsave("combinedV0.pdf", combined, width = 13, height = 5, dpi = 300)
```

